Linking microbial communities to ecosystem functions: what we can learn from genotype-phenotype mapping in organisms

### 2019-07-20 manuscript/main.tex
Revise Introduction

Population geneticists trying to make robust associations between genetic loci
and phenotype must consider several points when designing a genome-wide
association study. First is
to identify the precise trait expected to be influenced by genetic variation. 
Then, to choose a population that exhibits
substantial variation in the trait of interest. This population should also be
otherwise relatively homogeneous to avoid other factors that might influence the
trait. After sampling, statistical methods should take into account population
stratification due to shared ancestry or local selection that can lead to spurious
associations between genes and traits. For example, both human height and
genotype vary systematically across Europe in a latitudinal gradient, which can
lead to spurious genotype-phenotype associations \citep{novembre2008}.

Positive associations between genes and
phenotypes can arise because of true associations, through linkage
disequilibrium with true genes, or through admixture \citep{lander1994}.
Microbial structure-function connections could pose these same problems. There
could be true associations between a microorganism and the ecosystem function
performed by that taxon, or there could be systematic
co-occurrence patterns between one taxon and a causal taxon,
or there could be admixture due to shared history or selection within
a subset of ecosystems. This is true for 16S amplicons and for
genes in metagenomic (or other -omic) datsets. 
Robust structure-function associations rely on
controlling for the underlying geographic and ecological covariance.

Organisms exhibit population differentiation due to geographic and ecological
processes \citep{wright1943}. Neutral processes such as local genetic drift and
reduced gene flow due to geographic separation can lead to population
differentiation. In addition, environmental distance can lead to selection
against migrants \citep{hendry2004} and local selection on endemic populations
to generate variation. Microorganisms also exhibit biogeographic patterns due to
ecological selection and geographic separation \citep{martiny2006}. If dispersal
is non-limiting then the most productive species should always be present and
composition should not matter for local ecosystem function. However, if important
functional groups are dispersal-limited, lost to drift, or are unable to
colonize a site, then there could be important functional consequences of
community differentiation due to geographic or ecological isolation.

There are a number of ways that environmental or spatial distance could obscure
microbial BEF relationships. For example, historical effects of geographic
isolation such as dispersal limitation could lead to spurious associations
between the abundance of taxa and the rate of ecosystem function. Community
stratification where taxa co-occur due to historical effects or local selection.
Environmental selection can also shape microbial community
structure. For example, pH is known to regulate both broad biodiversity patterns for
microbial communities \citep{fierer2006} and methane cycling dynamics
\citep{ye2012}. This could lead to microbial taxa that are correlated with
methane cycling, but are mechanistically unrelated. In fact, if the rate of
ecosystem function is high in one set of ecosystems, function will be
associated with any taxa that are abundant in those ecosystems. 
In order to understand the biological basis of ecosystem function, we need to
separate the close association between environmental variation, community
composition, and ecosystem function in order to identify what aspects of
microbial biodiversity determine the rate of ecosystem function.

Successfully connecting microbial community structure to ecosystem function
depends on being precise about the variation we want to explain. The first and
often overlooked step is choosing a
clearly defined function that exhibits considerable variation across ecosystems. 
Next, be intentional about choosing the
population to be sampled to maximize the amount of variation. For example, how
might ecosystems that are otherwise quite similar vary in function?
Alternatively, how might one ecosystem change in response to a single
environmental change? This could be achieved by sampling across similar ecosystems 
or by imposing experimental conditions that minimize environmental variation or
vary only abiotic factors of interest. 
It is important to take an agnostic approach towards identifying taxa that are
indicators of ecosystem function. It is also important to remember that any
relationships identified in this manner are provisional until further tests are
performed. The work is not done until the organism has been manipulated through
addition, knockout, or variation in abundance.

In this paper we propose integrating microbial biodiversity-ecosystem function
research with ideas and techniques from population genetics to advance our
understanding of the role of microbial commuities in the functioning of
ecosystems. Perhaps the most important part of that discussion is to define the
question carefully. In microbial ecology
structure-function research is focused on answering the question: what is the
unique contribution of the microbial community to ecosystem function? This is of
great interest for two reasons. First, because it addresses what is the singular
contribution of biodiversity to the functioning of ecosystems. Second, on a
practical level, because answering that question will be essential to
incorporate a microbial component into ecosystem models.

Is variation in microbial community
structure associated with differences in ecosystem function across space,
time, and environments?

>>>>

Population geneticists make robust associations between genetic loci
and phenotypes by carefully considering study design when performing a genome-wide
association study. Studies should consider that positive associations between genes and
phenotypes can arise because of true associations, through linkage
disequilibrium with true genes, or through admixture where traits at high
frequency in a sub-population will be associated with any allele more common in
that group \citep{lander1994}. A common example of admixture is that human height and
genotype vary systematically across Europe in a latitudinal gradient, which can
lead to spurious genotype-phenotype associations \citep{novembre2008}. To
correct for this, it is
imporant to sample from a population that exhibits variation in the trait of
interest, but is otherwise relatively homogeneous
to avoid spurious associations due to admixture. In addition, it is important to standardize phenotyping
criteria to provide a narrowly defined trait. While certain genetic diseases are
controlled by only one or a few genetic loci, common diseases tend to be
controlled by many common variants of small effect. To detect these
associations requires large sample sizes for association tests. After
sampling the genotype and phenotype of this population, association tests should
adjust for population stratification using covariates, which are commonly
variables like sex and age as well as shared ancestry based on
genotypic similarity. Finally, after comparing thousands or millions of genetic
loci, investigators should correct for multiple comparisons.

Studies attempting to link microbial community structure to ecosystem function
could pose similar challenges. In this case, individual taxa are analogous to
genetic loci and their abundance or presence can be association with the rate of 
ecosystem function. There
could be true associations between a microorganism and the ecosystem function if
that organism is involved in the metabolic pathway producing that function.
Alternatively, there 
could be systematic co-occurrence patterns between a taxon and ecosystem
function due to admixture if that organism tends to reside in ecosystems with
high rates of ecosystem function.

Microbial admixture could arise through a
number of ecological mechanisms illustrated in conceptual diagram 1. For
example, an environmental condition such as pH could contribute to a high rate
of ecosystem function and select for a taxon even if that taxon is unrelated to
function. An association could also arise if an organism disperses from a
low-rate ecosystem to an ecosystem with a high rate of ecosystem function
through source-sink dynamics. In fact, if the rate of
ecosystem function is high in one set of ecosystems, function will be
associated with any taxa that are abundant in those ecosystems.
In order to understand the biological basis of ecosystem function, we need to
separate the close association between environmental variation, community
composition, and ecosystem function in order to identify what aspects of
microbial biodiversity determine the rate of ecosystem function.

%Microorganisms exhibit biogeographic patterns due to
%ecological selection and geographic separation \citep{martiny2006}. 
%Organisms exhibit population differentiation due to geographic and ecological
%processes \citep{wright1943}. Neutral processes such as local genetic drift and
%reduced gene flow due to geographic separation can lead to population
%differentiation. In addition, environmental distance can lead to selection
%against migrants \citep{hendry2004} and local selection on endemic populations
%to generate variation.

The same study design decisions used in GWAS could be applied to microbial
studies of structure-function connections. First, it is important to choose a 
subset of ecosystems that are relatively homogeneous in their conditions other
than that function of interest. This could be achieved by sampling across
similar types of ecosystems or by imposing conditions that minimize
environmental varition or vary only abiotic factors of interest. Next, it is important to provide a narrow
definition of function and a consistant measurement approach across studies.
Also, some functions are putatively controlled by one or a few taxa
\citep{mccalley2014}, but most functions are likely controlled by many
interacting microorganisms. To detect these associations depends on a large
sample size. After collecting samples, the underlying sample structure should be
controlled for by including environmental and spatial covariates and by
accounting for population structure. Finally, statistical approaches to adjust
for multiple comparisons should be applied. This approach could be applied to 16S amplicons or
genes in metagenomic (or other -omic) datasets.

%clearly defined function that exhibits considerable variation across ecosystems. 
%Next, be intentional about choosing the
%population to be sampled to maximize the amount of variation. For example, how
%might ecosystems that are otherwise quite similar vary in function?
%Alternatively, how might one ecosystem change in response to a single
%environmental change?
%It is important to take an agnostic approach towards identifying taxa that are
%indicators of ecosystem function. It is also important to remember that any
%relationships identified in this manner are provisional until further tests are
%performed. The work is not done until the organism has been manipulated through
%addition, knockout, or variation in abundance.

In this paper we propose integrating microbial biodiversity-ecosystem function
research with ideas and techniques from population genetics to advance our
understanding of the role of microbial communities in the functioning of
ecosystems. Perhaps the most important part of that discussion is to define the
question carefully. Microbial structure-function research should focus on answering the question: what is the
unique contribution of the microbial community to ecosystem function independent
of the environment? This is of
great interest for two reasons. First, because it addresses what is the singular
contribution of biodiversity to the functioning of ecosystems. Second, on a
practical level, because answering that question will be essential to
incorporate a microbial component into ecosystem models. To address that
question, we demonstrate that one ecosystem function, high-affinity methane
oxidation, is significantly association with variation in microbial community
composition. We also show that three components of population stratification
contribute significantly to variation in ecosystem function. Finally, we
demonstrate one approach to identifying the component of microbial
community that contributes to ecosystem function use an association test that
controls for these three
covariates and find four microbial taxa that are significantly associated with
ecosytsem function.
